\chapter{Bioliography}
\begin{flushleft}
\item{[1]} Dursun, M., Karsak, E.E., Karadayi, M.A., 2011a. Assessment of health-care waste
treatment alternatives using fuzzy multi-criteria decision making approaches.
Resour. Conserv. Recycl. 57, 98e107, 0.
\item{[2]} Dursun, M., Karsak, E.E., Karadayi, M.A., 2011b. A fuzzy multi-criteria group decision
making framework for evaluating health-care waste disposal alternatives.
Expert Syst. Appl. 38 (9), 11453e11462.
\item{[3]} Mishra, Arunodaya Raj, et al. "A novel EDAS approach on intuitionistic fuzzy set for assessment of health-care waste disposal technology using new parametric divergence measures." Journal of cleaner Production 272 (2020): 122807.
\item{[4]} Liu, Hu-Chen, Jing Wu, and Ping Li. "Assessment of health-care waste disposal methods using a VIKOR-based fuzzy multi-criteria decision making method." Waste management 33.12 (2013): 2744-2751.
\item{[5]} Zadeh LA (1965) Fuzzy sets. Inf Control 8(3):338–353
\item{[6]} Atanassov, Krassimir T., and Krassimir T. Atanassov. Intuitionistic fuzzy sets. Physica-Verlag HD, 1999.
\item{[7]} Alkan, Nurşah, and Cengiz Kahraman. "Evaluation of government strategies against COVID-19 pandemic using q-rung orthopair fuzzy TOPSIS method." Applied Soft Computing 110 (2021): 107653.
\item{[8]} Rouyendegh, Babak Daneshvar, Abdullah Yildizbasi, and Pelin Üstünyer. "Intuitionistic fuzzy TOPSIS method for green supplier selection problem." Soft Computing 24 (2020): 2215-2228.
\item{[9]}Hwang CL, Yoon K (1981) Multiple attribute decision making.
Methods and Applications Springer, New York
\item{[10]}Morrissey, A.J., Browne, J., 2004. Waste management models and their application
to sustainable waste management. Waste Manag. 24 (3), 297e308.
\item{[11]} Senapati, Tapan, et al. "Intuitionistic fuzzy geometric aggregation operators in the framework of Aczel-Alsina triangular norms and their application to multiple attribute decision making." Expert Systems with Applications 212 (2023): 118832.
\item{[12]}Liu, H.-C., Wu, J., Li, P., 2013. Assessment of health-care waste disposal methods
using a VIKOR-based fuzzy multi-criteria decision making method. Waste
Manag. 33 (12), 2744e2751.
\item{[13]}Z. Liu, P. Liu, X. Liang, Multiple attribute decision-making method for
dealing with heterogeneous relationship among attributes and unknown
attribute weight information under q-rung orthopair fuzzy environment,
Int. J. Intell. Syst. 33 (9) (2018) 1900–1928.
%\item{[14]} D. Liu, X. Chen, D. Peng, Some cosine similarity measures and distance
%measures between q-rung orthopair fuzzy sets, Int. J. Intell. Syst. 34 (7)(2019) 1572–1587.

\end{flushleft}